Left-linear rules in adhesive categories

### 2024-02-13 icalp/fusione.tex

\usepackage{pgf}
\usepackage{tikz}
\usepackage{tikz-cd}
\usetikzlibrary{arrows,shapes,snakes,automata,backgrounds,petri,fit,positioning}
\tikzstyle{node}=[circle, draw=black, minimum size=1mm]
\tikzstyle{trans}=[font=\scriptsize]
\tikzstyle{lab}=[font=\small]

@@

% \todo[inline]{
% This fails for left-linear rules where, as observed
% in~\cite{baldan2017domains}, for graph rewriting, the possibility of
% merging parts of the state leads to a sort of disjunctive causality,
% i.e., a rewriting step, say $r_c$ can be enabled by two different
% rewriting steps $r_a$ and $r_b$, in a way that in a derivation
% $r_a r_b r_c$ steps $r_b$ and $r_c$ are independent since the presence
% of $r_a$ in the past is sufficient for enabling $r_a$, while in
% $r_b r_c r_a$ , the steps $r_b$ and $r_c$ are not independent.
% }

% \todo[inline]{
% DA RIUSARE\\
% Firstly, Church-Rosser theorem does not work out of the box. In
% particular, sequential independence between rewriting steps~\cite{}
% expressed in terms of the existence of a so-called independence pair
% is a canonical and well-understood notion of independence in the
% linear case. \todo{P: Too technical? Forse qui e' piu' esplicativo un
%   esempio} Whenever two steps are sequential independent
% \begin{itemize}
% \item they can be switched and 
% \item the result of the switch is uniquely determined.
% \end{itemize}
% Here we observe that, in general, both property can fail: sequential
% independent steps could be not switchable and different independent
% pairs could exist leading to different switches of the same rewriting
% steps.  A refined notion of independence needs to be considered to ensure
% that independent rewriting steps can be switched, which is based in the concept
% of fillers. The problem was identified
% in~\cite{baldan2011adhesivity}, where a subclass of adhesive
% categories was identified where independence as induced by a filler
% coincides with sequential independence (as induced by an independence
% pair). Here it is extended to $\mathcal{M}$-adhesive categories.
% }

%   First of all, we will provide a refined notion of independence,
%   which is based in the concept of fillers.

% Then, we move to restrict the class of admissible left-linear rewriting systems,
% always working at the level of $\mathcal{M}$-adhesive categories.
% Unfortunately, the most natural idea, put forward in~\cite{baldan2017domains}, 
% of limiting sequential independence to those cases in which the switch is possible
% and uniquely determined does fall short.

@@

  We will say that two derivations $\der{D}, \der{E}:G\Mapsto H$ are \emph{switch
    equivalent} if there is a switching sequence, possibly empty, $\{\der{D}_{k}\}_{k=0}^m$ such that
      \[\der{D}\equiv_a \der{D}_0 \shift{\transp{i_1}} \der{D}_1 \shift{\transp{i_2}} \ldots \shift{\transp{i_m}}
    \der{D}_m\equiv _a \der{E}\]
\end{definition}

@@

\subsection{Switch equivalence via consistent permutations} \label{subsec:perm}

@@

	A \emph{consistent} permutation between two derivations $\der{D}=\{\dder{D}_i\}_{i=0}^n$ and
	$\der{D}'=\{\dder{D}'_i\}_{i=0}^n$ with associated sequence of rules
	$r(\der{D})=\{\rho_i\}_{i=0}^n$ is a permutation
	$\sigma\colon [0,n]\to [0,n]$ such that, for every $i\in [0,n]$,
	$\rho_i=\rho'_{\sigma(i)}$ and, moreover, there exists a
	\emph{mediating isomorphism}
	$\xi_\sigma\colon \tpro{D} \to \lpro \der{D}' \rpro$ fitting in the
	following diagrams, where $m_i, m'_i, h_i$ and $h'_i$ are,
	respectively, the matches and back-matches of $\dder{D}_i$ and
	$\dder{D}'_i$

	\[
	\xymatrix{
		G_0 \ar[r]^{\iota'_{G'_0}} \ar[d]_{\iota_{G_0}}& \lpro \der{D}' \rpro &L_i\ar[r]^{m_i} \ar[d]_{m'_{\sigma(i)}}& G_i \ar[r]^{\iota_{G_i}}
		&\tpro{D} \ar[d]^{\xi_\sigma} & R_i \ar[r]^{h_i}
		\ar[d]_{h'_{\sigma(i)}}& G_{i+1} \ar[r]^{\iota_{G_{i+1}}}
		&\tpro{D} \ar[d]^{\xi_\sigma} \\\tpro{D} \ar[ur]_{\xi_\sigma}&&G'_{\sigma(i)}
		\ar[rr]_{\iota'_{G'_{\sigma(i)}}}&& \lpro \der{D}' \rpro&
		G'_{\sigma(i)+1} \ar[rr]_{\iota'_{G'_{\sigma(i)+1}}}&& \lpro
		\der{D}' \rpro}
	\]
\end{definition}

@@

\subsection{A canonical form for  switch equivalences}
\label{subsec:canonical}

@@

\begin{example}
  Consider the rewriting system used as running example (rules in
  Fig.~\ref{fi:rules}. Assume to add rules $\rho_3$ and $\rho_4$ as
  depicted in Fig.~\ref{fi:rules-cas}. Consider the derivation
  $\der{F}$ in Fig.~\ref{fi:derF}. Observe that rule $\rho_4$ to be
  applied needs a self loop on node $12$. This can be generated either
  by $\rho_3$ or $\rho_1$ merging nodes $1$ and $2$, hence at least
  one of them must precede the application of $\rho_4$. Indeed, in $\der{E}$ it can be easily seen that the second and third step applying $\rho_1$ and $\rho_3$ are independent. However, we could switch twice the application of $\rho_3$, bringing it in the last position, obtaining a derivation $\der{F}'$ (see {\color{red} DA FINIRE, se ci sta}), where the applications of  $\rho_1$ and $\rho_3$ are no longer independent.

@@

  \begin{figure}
    \input{Figure/derF.tex}    
    \caption{A derivation $\der{F}$.} 
    \label{fi:derF}
  \end{figure}
\end{example}

@@

While in the case of linear rewriting systems inversion can be
performed in any order, this is no longer true for left-linear
rewriting systems.\todo{P: counter example} Still, we can identify a
canonical form for switching derivations.

@@

\subsubsection{Proofs for \Cref{subsec:perm}}

@@

\begin{corollary}
	\label{cor:ele}
	Let $\der{D}=\{\dder{D}_{i}\}_{i=0}^n$ be a derivation between $G$
	and $H$. Let $j$ and $k$ be two indexes less or equal than $n+1$ and
	suppose that $j< k$.  Consider two arrows $a\colon T\to G_j$,
	$b\colon T\to G_k$. If $\iota_{G_j}\circ a = \iota_{G_k}\circ b$,
	then there exist a unique arrow $c\colon T\to D_j $ such
	that \[f_j\circ c = a\qquad \iota_{D_j}\circ c =\iota_{G_k}\circ b\]
\end{corollary}
\begin{lemma}\label{lem:idevero}
	Let $\der{D}=\{\dder{D}_{i}\}_{i=0}^n$ and $\der{D}'=\{\dder{D}'_i\}_{i=0}^n$ be two non-empty derivations in a linear DPO-rewriting system $(\X, \R)$ . Let  $\sigma:[0,n-1]\to [0,n-1]$  is a consistent permutation and suppose that the set 
	\[E_{\sigma}=\{i\in [0,n-1]\mid \sigma(i)=i\}\]
	is non empty with maximum $k$. Then $\{\dder{D}_i\}_{i=0}^k$ and $\{\der{D}'_i\}_{i=0}^k$ are abstraction equivalent.  
\end{lemma}

\begin{proof}We construct are going to construct by induction on $i\in [0,n]$ an abstraction equivalence such that, for every $i$, $\iota'_{G'_{i+1}}\circ \phi_{G_{i+1}}=\xi_{\id{[0,n]}}\circ \iota_{G_{i+1}}$.

@@

\begin{corollary}\label{cor:coswap}
Let $\der{D}$ and $\der{E}$ be two derivations such that $\der{D} \shift{\transp{i}} \der{E}$, then the transposition $(i,i+1):[0,\lgh{\der{D}}-1] \to [0,\lgh{\der{D}}-1] $ is a consistent permutation.
\end{corollary}

@@

\begin{lemma}
	Let $\der{D}$ and $\der{D}$ be two derivations in a linear rewriting system $(\X, \R)$. Suppose that a consistent permutation $\sigma:[0, \lgh(\der{D})-1]\to [0, \lgh(\der{D})-1]$ between them is given. If $i\in [0, \lgh(\der{D})-1]$ is the smallest index such that $\transp{i}$ is an inversion for $\sigma$, then the direct derivations $\dder{D}_i$ and $\dder{D}_{i+1}$ are sequentially independent.
\end{lemma}
\begin{proof}
	Let us start with the diagram below
	  \[\xymatrix@C=15pt{L_i \ar[d]_{m_i}&& K_i
		\ar[d]_{k_i}\ar@{>->}[ll]_{l_i} \ar@{>->}[r]^{r_i} & R_i
		\ar@{.>}@/^.35cm/[drrr]|(.3)\hole_(.4){j_0}
		\ar[dr]|(.3)\hole_{h_i} && L_{i+1} \ar@{.>}@/_.4cm/[dlll]^(.4){j_1}
		\ar[dl]|(.3)\hole^{m_{i+1}}& K_{i+1} \ar[d]^{k_{i+1}}\ar@{>->}[l]_{l_{i+1}}
		\ar@{>->}[rr]^{r_{i+1}} && R_{i+1} \ar[d]^{h_{i+1}}\\G_i && \ar@{>->}[ll]^{f_i}
		D_i \ar@{>->}[rr]_{g_i}&& G_{i+1} && \ar@{>->}[ll]^{f_{i+1}} D_{i+1}\ar@{>->}[rr]_{g_{i+1}}&& G_{i+2}}\]
	\parbox{9cm}{Notice that $\sigma$ restrict to the identity $[0, i-1]\to [0,i-1]$, which remains consistent. In particular, there exists an isomorphism $\phi_{i}\colon G_i\to G'_i$ such that the square on the right commutes.}
	\parbox{3cm}{\vspace{-.5cm}$\xymatrix{G_i \ar[d]_{\iota_{G_i}}\ar@/^.4cm/[rr]^{\phi_{G_i}}& L_0  \ar[l]^{m_i}\ar[r]_{m'_{i}} & G'_{i} \ar[d]^{\iota'_{G'_i}}\\ \tpro{D} \ar[rr]_{\xi_\sigma} & &\lpro\der{D}'\rpro  }$}

@@

\begin{remark}\label{rem:salvavita}
	Before proceeding further, it is worth to point out the following fact about inversions.
\end{remark}
\thmcoswitch*
\label{thmswitch-proof}
\begin{proof}$(1\Rightarrow 2)$ This follows from \Cref{rem:abs,prop:coswitch}.

\smallskip \noindent $(2\Rightarrow 1)$ We proceed by induction
\end{proof}

>>>>

\usepackage{pgf}
\usepackage{tikz}
\usepackage{tikz-cd}
\usetikzlibrary{arrows,shapes,snakes,automata,backgrounds,petri,fit,positioning}
\tikzstyle{node}=[circle, draw=black, minimum size=1mm, inner sep=1.5pt]
\tikzstyle{trans}=[font=\scriptsize]
\tikzstyle{lab}=[font=\small]

@@

@@

  We will say that two derivations $\der{D}, \der{E}:G\Mapsto H$ are \emph{switch
    equivalent} if there is a switching sequence, possibly empty, $\{\der{D}_{k}\}_{k=0}^m$ such that
      \[\der{D}\equiv_a \der{D}_0 \shift{\transp{i_1}} \der{D}_1 \shift{\transp{i_2}} \ldots \shift{\transp{i_m}}
    \der{D}_m\equiv _a \der{E}\]
    We will denote this relation by $\der{D}\equiv^{sh} \der{E}$. 
\end{definition}

@@

\iffalse 
\subsection{Switch equivalence via consistent permutations} \label{subsec:perm}

@@

	A \emph{consistent} permutation between two derivations $\der{D}=\{\dder{D}_i\}_{i=0}^n$ and
	$\der{D}'=\{\dder{D}'_i\}_{i=0}^n$ with associated sequence of rules
	$r(\der{D})=\{\rho_i\}_{i=0}^n$ is a permutation
	$\sigma\colon [0,n]\to [0,n]$ such that, for every $i\in [0,n]$,
	$\rho_i=\rho'_{\sigma(i)}$ and, moreover, there exists a
	\emph{mediating isomorphism}
	$\xi_\sigma\colon \tpro{D} \to \lpro \der{D}' \rpro$ fitting in the
	following diagrams, where $m_i, m'_i, h_i$ and $h'_i$ are,
	respectively, the matches and back-matches of $\dder{D}_i$ and
	$\dder{D}'_i$.
	\[
	\xymatrix{
		G_0 \ar[r]^{\iota'_{G'_0}} \ar[d]_{\iota_{G_0}}& \lpro \der{D}' \rpro &L_i\ar[r]^{m_i} \ar[d]_{m'_{\sigma(i)}}& G_i \ar[r]^{\iota_{G_i}}
		&\tpro{D} \ar[d]^{\xi_\sigma} & R_i \ar[r]^{h_i}
		\ar[d]_{h'_{\sigma(i)}}& G_{i+1} \ar[r]^{\iota_{G_{i+1}}}
		&\tpro{D} \ar[d]^{\xi_\sigma} \\\tpro{D} \ar[ur]_{\xi_\sigma}&&G'_{\sigma(i)}
		\ar[rr]_{\iota'_{G'_{\sigma(i)}}}&& \lpro \der{D}' \rpro&
		G'_{\sigma(i)+1} \ar[rr]_{\iota'_{G'_{\sigma(i)+1}}}&& \lpro
		\der{D}' \rpro}
	\]
\end{definition}

@@

\fi 
\subsection{A canonical form for  switch equivalences}
\label{subsec:canonical}

@@

\begin{example}
  \label{ex:non-global}
  Consider the rewriting system used as running example (rules in
  Fig.~\ref{fi:rules}. Assume to add rules $\rho_3$ and $\rho_4$ as
  depicted in Fig.~\ref{fi:rules-cas}. Consider the derivation
  $\der{F}$ in Fig.~\ref{fi:derF}. Observe that rule $\rho_4$ to be
  applied needs a self loop on node $12$. This can be generated either
  by $\rho_3$ or $\rho_1$ merging nodes $1$ and $2$, hence at least
  one of them must precede the application of $\rho_4$. Indeed, in $\der{E}$ it can be easily seen that the second and third step applying $\rho_1$ and $\rho_3$ are independent. However, we could switch twice the application of $\rho_3$, bringing it in the last position, obtaining a derivation $\der{F}'$ as deipcted in Fig.~\ref{fi:derF1}, where the applications of  $\rho_1$ and $\rho_3$ are no longer independent.

@@

  \begin{figure}
    \input{Figure/derF.tex}    
    \caption{A derivation $\der{F}$.} 
    \label{fi:derF}
  \end{figure}

  \begin{figure}
    \input{Figure/derF1.tex}    
    \caption{A derivation $\der{F}'$.} 
    \label{fi:derF1}
  \end{figure}
\end{example}

@@

While in the case of linear rewriting systems inversion can be
performed in any order, this is no longer true for left-linear
rewriting systems due to the fact that, as already observed globality of independence holds in a weak form. For instance, if we get back to Example~\ref{ex:non-global} and consider derivations $\der{F}$ in Fig.~\ref{fi:derF}, the last two steps applying rules $\rho_1$ and $\rho_4$ can be switched, thus leading to a derivation, call it $\der{F}''$,  applying $\rho_3, \rho_4, \rho_1$. Such $\der{F}$ is switch equivalent to $\der{F}'$ in Fig.~\ref{fi:derF1}, hence starting from $\der{F}''$ we can obtain $\der{F}'$ via a switching sequence, but we cannot start by switching $\rho_1$ and $\rho_4$, despite the fact that this is an inversion.

Still, we can identify a
canonical form for switching derivations.

@@

\iffalse 
\subsubsection{Proofs for \Cref{subsec:perm}}

@@

\begin{lemma}\label{lem:idevero}
	Let $\der{D}=\{\dder{D}_{i}\}_{i=0}^n$ and $\der{D}'=\{\dder{D}'_i\}_{i=0}^n$ be two non-empty derivations in a linear DPO-rewriting system $(\X, \R)$ . Let  $\sigma:[0,n-1]\to [0,n-1]$  is a consistent permutation and suppose that the set 
	\[E_{\sigma}=\{i\in [0,n-1]\mid \sigma(i)=i\}\]
	is non empty with maximum $s_\sigma$. Then $\{\dder{D}_i\}_{i=0}^{s_\sigma}$ and $\{\der{D}'_i\}_{i=0}^{s_\sigma}$ are abstraction equivalent.  
\end{lemma}

\begin{proof}We construct are going to construct by induction on $i\in [0,s_\sigma]$ an abstraction equivalence such that, for every $i$, $\iota'_{G'_{i+1}}\circ \phi_{G_{i+1}}=\xi_{\id{[0,n]}}\circ \iota_{G_{i+1}}$.

@@

\begin{corollary}\label{cor:coswap}
Let $\der{D}$ and $\der{E}$ be two derivations such that $\der{D} \shift{\transp{i}} \der{E}$, then the transposition $(i,i+1):[0,\lgh(\der{D})-1] \to [0,\lgh(\der{D})-1] $ is a consistent permutation.
\end{corollary}

@@

\iffalse 
\begin{remark}\label{rem:salvavita}
Before proceeding further, it is worth to point out the following fact about inversions.  Take a permutation $\sigma\colon [0,1]\to [0,1]$ and define 
\[\tnv{\sigma}:=\{(i,j)\in \inv{\sigma}\mid j=i+1\}\]

We still can prove that $\sigma$ is the identity if and only if $\tnv{\sigma}=\empty$. Indeed, if $(i,j)$ is an inversion for$\sigma$, then we have two case:
\begin{itemize}
	\item for every $h,k\in[i,j-1]$, $(h,k)$ is not an inversion, thus $(j-1, j)$ is an inversion belonging to $\tnv{\sigma}$; 
	\item there exists an inversion $(h,k)$ made by elements of $[i,j-1]$: then we can repeat the same argument and conclude proceeding by induction on $k-h$.
\end{itemize}

Now, let $\sigma$ be a permutation different from the identity, recall that $E_{\sigma}$ is the set all all $i$ such that $\sigma(i)=i$, then we have:
\begin{itemize}
	\item if $E_{\sigma}=\emptyset$ then the minimum $i$ such that $(i-1,i)$ belongs to $\tnv{\sigma}$ is $\sigma^{-1}(0)$. On the one hand,  $\sigma(\sigma^{-1}(0)-1)$ cannot be $0$ and thus $(\sigma^{-1}(0)-1, \sigma^{-1}(0))$ is in $\tnv{\sigma}$. On the other hand  
	\item 
\end{itemize}

let $k_{\sigma}$ the minimum $i$ such that $(i, i+1)$
\end{remark}
\fi

\begin{lemma}
	Let $\der{D}$ and $\der{D}$ be two derivations in a linear rewriting system $(\X, \R)$. Suppose that a consistent permutation $\sigma:[0, \lgh(\der{D})-1]\to [0, \lgh(\der{D})-1]$ between them is given. Then the following hold true:
	\begin{enumerate}
		\item if $E_{\sigma} = \emptyset$, so that $\sigma(0)>0$ then the direct derivations $\dder{D}_{\sigma^{-1}(0)-1}$ and $\dder{D}_{\sigma^{-1}(0)}$ are sequentially independent;
		\item if $E_{\sigma}\neq \emptyset$ and $s_\sigma$ is its maximum, then $\dder{D}_{\sigma^{-1}(s_\sigma)-1}$ and $\dder{D}_{\sigma^{-1}(s_\sigma)}$ are sequentially independent.
	\end{enumerate}
\end{lemma}
\begin{proof}
	In both cases, we start with the diagram below, where $t=\sigma^{-1}(0)-1$ or $t=\sigma{^-1}(s_\sigma)-1$
	\[\xymatrix@C=15pt{L_t \ar[d]_{m_t}&& K_t
		\ar[d]_{k_t}\ar@{>->}[ll]_{l_t} \ar@{>->}[r]^{r_t} & R_t
		\ar@{.>}@/^.35cm/[drrr]|(.3)\hole_(.4){i_0}
		\ar[dr]|(.3)\hole_{h_t} && L_{t+1} \ar@{.>}@/_.4cm/[dlll]^(.4){i_1}
		\ar[dl]|(.3)\hole^{m_{t+1}}& K_{t+1} \ar[d]^{k_{t+1}}\ar@{>->}[l]_{l_{t+1}}
		\ar@{>->}[rr]^{r_{t+1}} && R_{t+1} \ar[d]^{h_{t+1}}\\G_t && \ar@{>->}[ll]^{f_t}
		D_t \ar@{>->}[rr]_{g_t}&& G_{t+1} && \ar@{>->}[ll]^{f_{t+1}} D_{t+1}\ar@{>->}[rr]_{g_{t+1}}&& G_{t+2}}\]

	\begin{enumerate}
		\item If $E_{\sigma}=\emptyset$, so that $t=\sigma^{-1}(0)-1$, then $\sigma(t+1)=0$. Thus
		\[\xi_\sigma\circ \iota_{G_0}\circ m_{0}=\iota'_{G_0}\circ m'_{}\]
		 Therefore we have the solid part of the diagram below. \[\xymatrix@C=30pt{D_{j} \ar[ddrr]^{\iota_{D_j}}
		 	\ar[r]^{g_j}\ar@{>->}[d]_{f_j}& G_{j+1}
		 	\ar[ddr]^{\iota_{G_{j+1}}}
		 	\ar[r]^-{\iota_{3, G_{j+1}}} & \lpro \der{D}^j_3\rpro \ar[dd]^{p_2} \\
		 	G_0\ar[drr]_{\iota_{G_{j}}} \ar@{>->}[d]_{\iota_{1,G_j}}\\ \lpro
		 	\der{D}^j_1 \rpro \ar[rr]_{p_1} &&\tpro{D} }\]

		 	and\Cref{cor:colim} yields $i_0\colon L_{t+1}\to$
		\item 
	\end{enumerate}
	Let us start with the diagram below
	\[\xymatrix@C=15pt{L_i \ar[d]_{m_i}&& K_i
		\ar[d]_{k_i}\ar@{>->}[ll]_{l_i} \ar@{>->}[r]^{r_i} & R_i
		\ar@{.>}@/^.35cm/[drrr]|(.3)\hole_(.4){j_0}
		\ar[dr]|(.3)\hole_{h_i} && L_{i+1} \ar@{.>}@/_.4cm/[dlll]^(.4){j_1}
		\ar[dl]|(.3)\hole^{m_{i+1}}& K_{i+1} \ar[d]^{k_{i+1}}\ar@{>->}[l]_{l_{i+1}}
		\ar@{>->}[rr]^{r_{i+1}} && R_{i+1} \ar[d]^{h_{i+1}}\\G_i && \ar@{>->}[ll]^{f_i}
		D_i \ar@{>->}[rr]_{g_i}&& G_{i+1} && \ar@{>->}[ll]^{f_{i+1}} D_{i+1}\ar@{>->}[rr]_{g_{i+1}}&& G_{i+2}}\]
	We have two cases.

	\smallskip \noindent $i=0$.

	\smallskip\noindent $i>0$

	\parbox{9cm}{Notice that $\sigma$ restrict to the identity $[0, i-1]\to [0,i-1]$, which remains consistent. In particular, there exists an isomorphism $\phi_{i}\colon G_i\to G'_i$ such that the square on the right commutes.}
	\parbox{3cm}{\vspace{-.5cm}$\xymatrix{G_i \ar[d]_{\iota_{G_i}}\ar@/^.4cm/[rr]^{\phi_{G_i}}& L_0  \ar[l]^{m_i}\ar[r]_{m'_{i}} & G'_{i} \ar[d]^{\iota'_{G'_i}}\\ \tpro{D} \ar[rr]_{\xi_\sigma} & &\lpro\der{D}'\rpro  }$}

@@

\begin{lemma}\label{lem:fix1}
	Let $\der{D}$ and $\der{D}'$ be two derivations in a linear rewriting system $(\X, \R)$. Suppose that a consistent permutation $\sigma:[0, \lgh(\der{D})-1]\to [0, \lgh(\der{D})-1]$ between them is given. Then there is a derivation $\der{E}$ and a switching sequence 
	  \[\der{D}\equiv_a \der{D}_0 \shift{\nu_1} \der{D}_1 \shift{\nu_2} \ldots \shift{\transp{\nu_m}}
	\der{D}_m = \der{E}\]
proving that  $\der{D}\equiv^{sh} \der{E}$ and such that $\sigma\circ \nu_{1,m}(0)=0$.
\end{lemma}
\begin{proof}
	contenuto...
\end{proof}

\begin{lemma}\label{lem:fix2}
	Let $\der{D}$ and $\der{D}'$ be two derivations in a linear rewriting system $(\X, \R)$. Suppose that a consistent permutation $\sigma:[0, \lgh(\der{D})-1]\to [0, \lgh(\der{D})-1]$ between them is given. Then there is a derivation $\der{E}$ and a switching sequence 
	\[\der{D}\equiv_a \der{D}_0 \shift{\nu_1} \der{D}_1 \shift{\nu_2} \ldots \shift{\transp{\nu_m}}
	\der{D}_m = \der{E}\]
	proving that  $\der{D}\equiv^{sh} \der{E}$ and such that $\sigma\circ \nu_{1,m}(0)=0$.
\end{lemma}
\begin{proof}
	contenuto...
\end{proof}

\thmcoswitch*
\label{thmswitch-proof}
\begin{proof}$(1\Rightarrow 2)$ This follows from \Cref{rem:abs,prop:coswitch}.

\smallskip \noindent $(2\Rightarrow 1)$ We proceed by induction
\end{proof}
\fi
